---
title: "Status of climate adaptation in the Australian Water Utilities: results from survey of higher level managers"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: '2'
  html_notebook:
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 2
  pdf_document:
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 2
---

<!--For automatic fig captioning and citation, see http://derekogle.com/fishR/2015-09-17-Figure-Table-Captions-in-Markdown-->

```{r echo=FALSE, results='hide'}
utility_name <- 'South East Water'
```

# Results for `r utility_name` and comparison to national data from other utilities surveyed

We conducted a survey, targeted at higher-level managers, with the primary aim of identifying officers that have produced climate adaptation information. In addition, we also wanted to capture the extent to which climate adaptation is incorporated within decision making and the importance of climate change in the mindset of managers and within leadership. This is because the role of climate-adaptation within water utilities is significantly determined by higher-level managers and leaders. In this document, we present detailed results of this survey, comparing responses arising from `r utility_name` with the other Australian utilities surveyed. The utilities taking part in this survey were Water Corporation of Western Australia, South Australian Water Corporation, Melbourne Water, South East Water, City West Water, Yarra Valley Water, Coliban Water, Barwon Water, Sydney Water, and Hunter Valley Water.

```{r echo=FALSE, results='hide'}
library(readxl)
df <- read_excel('/home/a1091793/Documents/Projects/WaterRA Data driven Climate Adaptation/Survey/Results 23 Feb 2018/High level survey of climate adaptation v2 aggregated.xlsx')
df_utility <- df[df$WaterUtility == utility_name, ]
df_other_utilities <- df[df$WaterUtility != utility_name, ]
```

## Survey respondents
The following people responded to this survey:

```{r echo=FALSE, results='asis'}
cat(paste('-', df_utility$Name, " (", df_utility$CoreArea, ")"), sep="\n")
```

The following people were identified by the respondents for inclusion in the more detailed survey on developing climate adaptation information.

```{r echo=FALSE, results='asis'}
cat(paste('-', df_utility$ClimateOfficers[!is.na(df_utility$ClimateOfficers)]), sep="\n")
```

These people were also identified as being important within the climate adaptation process in the water utlity:

```{r echo=FALSE, results='asis'}
cat(paste('-',df_utility$OtherPeople[!is.na(df_utility$OtherPeople)]), sep="\n")
```

```{r echo=FALSE, results='hide'}
# Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

## Embedment of climate adaptation within water utilities

The WSAA guidelines provide suggestions on how climate adaptation can be embedded within a water utility focussing on the board, chief executive, top-level management, resourcing and lower-level roles. We asked respondents whether these suggestions are in place within the utilities they work for.  

### Leadership considers that climate change is a key risk that needs addressing from a whole-of-business perspective

Water utilities differ in their size, scope, organisation and business processes. Therefore, specific staffing and resourcing requirements cannot be generalised across water utilities. Saying that, general guidance can be provided, including that the leadership at the highest level acknowledges that climate change is a key risk and supports addressing this risk from a whole-of-business perspective. In accordance to this suggestion provided by the WSAA Guidelines, we asked whether this leadership existed at the board, chief executive office and senior management level.  

#### Leadership at the board level

The figure below (Fig. 1.) shows the level to which the water utility board considers climate adaptation as a key risk that needs addressing from a whole-of-business perpective, as stated by managers at their utility. As can be seen, managers tended to view the board as strongly supporting climate adaptation.  

```{r fig.width=7, fig.height=2, echo=FALSE}
library(captioner)
figs <- captioner(prefix="Figure")
library(grid)
grid.newpage()

board_support_v <- df$BoardSupport
board_support_v  <- factor(board_support_v, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
board_support_df <- as.data.frame(table(board_support_v))
library(ggplot2)
g <- ggplot(data=board_support_df, aes(x=board_support_v, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") +xlab("Strength of aggreement") + theme(axis.text.x = element_blank())  + ggtitle("National results"), vp = viewport(x=0, y=0, width = 0.6, height = 1, just = c('left', 'bottom')))

board_support_v_utility <- df_utility$BoardSupport
board_support_v_utility  <- factor(board_support_v_utility, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
board_support_df_utility <- as.data.frame(table(board_support_v_utility))
library(ggplot2)
g_utility <- ggplot(data=board_support_df_utility, aes(x=board_support_v_utility, y=Freq))
print(g_utility + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") +xlab("") + theme(axis.text.x = element_blank(), axis.text.y = element_blank()) + ggtitle(paste(utility_name, "results")), vp = viewport(x=0.6, y=0, width = 0.4, height = 1, just = c('left', 'bottom')))
```
```{r echo=FALSE, results='hide'}
figs(name="board_leadership", "Frequency of responses by water utility managers to the question: \"The board considers that climate change is a key risk that needs addressing from a whole-of-business perspective,\" given a rating scale between 'strongly disagree' to 'strongly agree.'")
```
`r figs("board_leadership")`


#### Leadershop at the chief executive officer level

Fig. 2, below, shows the level to which the chief executive officer considers climate adaptation as a key risk that needs addressing from a whole-of-business perpective, as stated by managers at their utility. As can be seen, managers tended to view the CEO as strongly supporting climate adaptation, with similar results here as for those regarding board level support.  

```{r fig.width=7, fig.height=2, echo=FALSE}
grid.newpage()

chief_executive_support_v <- df$ChiefExecutyiveSupport
chief_executive_support_v  <- factor(chief_executive_support_v, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
chief_executive_support_df <- as.data.frame(table(chief_executive_support_v))
g <- ggplot(data=chief_executive_support_df, aes(x=chief_executive_support_v, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") +xlab("Strength of aggreement") + theme(axis.text.x = element_blank())  + ggtitle("National results"), vp = viewport(x=0, y=0, width = 0.6, height = 1, just = c('left', 'bottom')))

chief_executive_support_v_utility <- df_utility$ChiefExecutyiveSupport
chief_executive_support_v_utility  <- factor(chief_executive_support_v_utility, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
chief_executive_support_df_utility <- as.data.frame(table(chief_executive_support_v_utility))
g <- ggplot(data=chief_executive_support_df_utility, aes(x=chief_executive_support_v_utility, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") +xlab("")  + theme(axis.text.x = element_blank(), axis.text.y = element_blank()) + ggtitle(paste(utility_name, "results")), vp = viewport(x=0.6, y=0, width = 0.4, height = 1, just = c('left', 'bottom')))
```
Fig. 2. Frequency of responses by water utility managers to the question: "The chief executive officer considers that climate change is a key risk that needs addressing from a whole-of-business perspective," given a rating scale between 'strongly disagree' to 'strongly agree.'  

#### leadership at the senior management level

Fig. 3, below, shows the level to which senior managers support climate adaptation. Again, climate adaptation is shown to be highly supported at senior management level, although the average agreement to the question 'Senior management considers that Climate change is a key risk that needs addressing from a whole-of-business perspective' was slightly lower than those provided for this question at the board and CEO level.  

```{r fig.width=7, fig.height=2, echo=FALSE} 
grid.newpage()
senior_management_support_v <- df$ChiefExecutyiveSupport
senior_management_support_v  <- factor(senior_management_support_v, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
senior_management_support_df <- as.data.frame(table(senior_management_support_v))
g <- ggplot(data=senior_management_support_df, aes(x=senior_management_support_v, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") +xlab("Strength of aggreement") + theme(axis.text.x = element_blank())  + ggtitle("National results"), vp = viewport(x=0, y=0, width = 0.6, height = 1, just = c('left', 'bottom')))

senior_management_support_v_utility <- df_utility$ChiefExecutyiveSupport
senior_management_support_v_utility  <- factor(senior_management_support_v_utility, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
senior_management_support_df_utility <- as.data.frame(table(senior_management_support_v_utility))
g <- ggplot(data=senior_management_support_df_utility, aes(x=senior_management_support_v_utility, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") +xlab("")  + theme(axis.text.x = element_blank(), axis.text.y = element_blank()) + ggtitle(paste(utility_name, "results")), vp = viewport(x=0.6, y=0, width = 0.4, height = 1, just = c('left', 'bottom')))
```
Fig. 3. Frequency of responses by water utility managers to the question: "Senior management considers that Climate change is a key risk that needs addressing from a whole-of-business perspective," given a rating scale between 'strongly disagree' to 'strongly agree.'  

### Resourcing adaptation

Resources are required for adaptation; and whether these are personnel, cash-funding, information, data, guidelines or the availability of computational resources, they all need to be provided.  In accordance to suggestions provided by the WSAA Guidelines, we asked a number of questions related to the resourcing of adaptation.  

#### Resources for the coordination of, and input into the adaptation process
As shown in Fig 4, below, resources are available for the coordination of, and input into the adaptation option. However, it is notable that the strength of response by senior managers regarding resourcing is lower than that for support from leadership, as given in Figures 1-3.  

```{r fig.width=7, fig.height=2, echo=FALSE} 
grid.newpage()
sufficient_resiources_v <- df$SufficientResources
sufficient_resiources_v  <- factor(sufficient_resiources_v, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
sufficient_resiources_df <- as.data.frame(table(sufficient_resiources_v))
g <- ggplot(data=sufficient_resiources_df, aes(x=sufficient_resiources_v, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") +xlab("Strength of aggreement") + theme(axis.text.x = element_blank())  + ggtitle("National results"), vp = viewport(x=0, y=0, width = 0.6, height = 1, just = c('left', 'bottom')))

sufficient_resiources_v_utility <- df_utility$SufficientResources
sufficient_resiources_v_utility  <- factor(sufficient_resiources_v_utility, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
sufficient_resiources_df_utility <- as.data.frame(table(sufficient_resiources_v_utility))
g <- ggplot(data=sufficient_resiources_df_utility, aes(x=sufficient_resiources_v_utility, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") +xlab("")  + theme(axis.text.x = element_blank(), axis.text.y = element_blank()) + ggtitle(paste(utility_name, "results")), vp = viewport(x=0.6, y=0, width = 0.4, height = 1, just = c('left', 'bottom')))
```
Fig. 4. Frequency of responses by water utility managers to the question: "Sufficient resources are provided for coordination and input into the adaptation process," given a rating scale between 'strongly disagree' to 'strongly agree.'  

### Senior management is committed to resourcing climate adaptation

```{r fig.width=7, fig.height=2, echo=FALSE} 
grid.newpage()
senior_management_resourcing_v <- df$SeniorManagementResourcing
senior_management_resourcing_v  <- factor(senior_management_resourcing_v, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
senior_management_resourcing_df <- as.data.frame(table(senior_management_resourcing_v))
g <- ggplot(data=senior_management_resourcing_df, aes(x=senior_management_resourcing_v, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") +xlab("Strength of aggreement") + theme(axis.text.x = element_blank())  + ggtitle("National results"), vp = viewport(x=0, y=0, width = 0.6, height = 1, just = c('left', 'bottom')))

senior_management_resourcing_v_utility <- df_utility$SeniorManagementResourcing
senior_management_resourcing_v_utility  <- factor(senior_management_resourcing_v_utility, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
senior_management_resourcing_df_utility <- as.data.frame(table(senior_management_resourcing_v_utility))
g <- ggplot(data=senior_management_resourcing_df_utility, aes(x=senior_management_resourcing_v_utility, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") +xlab("")  + theme(axis.text.x = element_blank(), axis.text.y = element_blank()) + ggtitle(paste(utility_name, "results")), vp = viewport(x=0.6, y=0, width = 0.4, height = 1, just = c('left', 'bottom')))
```
Fig. X. Frequency of responses by water utility managers to the question: "YYY," given a rating scale between 'strongly disagree' to 'strongly agree.'

### Resources still lacking for climate adaptation processes

```{r echo=FALSE, results='hide'}
responses <- df_other_utilities$ResourcesRequired
responses <- responses[!is.na(responses)]
responses_utility <- df_utility$ResourcesRequired
responses_utility <- responses_utility[!is.na(responses_utility)]
```


Responses from `r utility_name` were:

```{r echo=FALSE, results='asis'}
cat(paste('-', responses_utility), sep = '\n') 
```

Additional responses from those in other Australian utilities, included:

```{r echo=FALSE, results='asis'}
cat(paste('-', responses), sep = '\n') 
```


### A policy, position, strategy or direction-setting document provides clear, transparent and ongoing commitment to adaptation planning over the long-term at my water utility/organisation

```{r fig.width=7, fig.height=2, echo=FALSE} 
grid.newpage()
strategy_document_v <- df$StrategyDocument
strategy_document_v  <- factor(strategy_document_v, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
strategy_document_df <- as.data.frame(table(strategy_document_v))
g <- ggplot(data=strategy_document_df, aes(x=strategy_document_v, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") +xlab("Strength of aggreement") + theme(axis.text.x = element_blank())  + ggtitle("National results"), vp = viewport(x=0, y=0, width = 0.6, height = 1, just = c('left', 'bottom')))

strategy_document_v_utility <- df_utility$StrategyDocument
strategy_document_v_utility  <- factor(strategy_document_v_utility, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
strategy_document_df_utility <- as.data.frame(table(strategy_document_v_utility))
g <- ggplot(data=strategy_document_df_utility, aes(x=strategy_document_v_utility, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") +xlab("Strength of aggreement")  + theme(axis.text.x = element_blank(), axis.text.y = element_blank()) + ggtitle(paste(utility_name, "results")), vp = viewport(x=0.6, y=0, width = 0.4, height = 1, just = c('left', 'bottom')))
```
Fig. X. Frequency of responses by water utility managers to the question: "YYY," given a rating scale between 'strongly disagree' to 'strongly agree.'

### The CEO has sufficient time allocated to reviewing the implementation and success of adaptation measures

```{r fig.width=7, fig.height=2, echo=FALSE} 
grid.newpage()
ceo_time_v <- df$CEOAllocatedTime
ceo_time_v  <- factor(ceo_time_v, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
ceo_time_df <- as.data.frame(table(ceo_time_v))
g <- ggplot(data=ceo_time_df, aes(x=ceo_time_v, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") +xlab("Strength of aggreement") + theme(axis.text.x = element_blank())  + ggtitle("National results"), vp = viewport(x=0, y=0, width = 0.6, height = 1, just = c('left', 'bottom')))

ceo_time_v_utility <- df_utility$CEOAllocatedTime
ceo_time_v_utility  <- factor(ceo_time_v_utility, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
ceo_time_df_utility <- as.data.frame(table(ceo_time_v_utility))
g <- ggplot(data=ceo_time_df_utility, aes(x=ceo_time_v_utility, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") +xlab("")  + theme(axis.text.x = element_blank(), axis.text.y = element_blank()) + ggtitle(paste(utility_name, "results")), vp = viewport(x=0.6, y=0, width = 0.4, height = 1, just = c('left', 'bottom')))
```
Fig. X. Frequency of responses by water utility managers to the question: "YYY," given a rating scale between 'strongly disagree' to 'strongly agree.'

### The board includes someone I would regard as a champion for climate change adaptation

```{r fig.width=7, fig.height=2, echo=FALSE} 
grid.newpage()
champion_v <- df$ChampionOnBoard
champion_v  <- factor(champion_v, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
champion_df <- as.data.frame(table(champion_v))
g <- ggplot(data=champion_df, aes(x=champion_v, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("Strength of aggreement") + theme(axis.text.x = element_blank())  + ggtitle("National results"), vp = viewport(x=0, y=0, width = 0.6, height = 1, just = c('left', 'bottom')))

champion_v_utility <- df_utility$ChampionOnBoard
champion_v_utility  <- factor(champion_v_utility, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
champion_df_utility <- as.data.frame(table(champion_v_utility))
g <- ggplot(data=champion_df_utility, aes(x=champion_v_utility, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("")  + theme(axis.text.x = element_blank(), axis.text.y = element_blank()) + ggtitle(paste(utility_name, "results")), vp = viewport(x=0.6, y=0, width = 0.4, height = 1, just = c('left', 'bottom')))
```
Fig. X. Frequency of responses by water utility managers to the question: "YYY," given a rating scale between 'strongly disagree' to 'strongly agree.'

### There is efficient coordination of climate adaptation activity at the senior management level

```{r fig.width=7, fig.height=2, echo=FALSE} 
grid.newpage()
coordination_v <- df$Coordination
coordination_v  <- factor(coordination_v, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
coordination_df <- as.data.frame(table(coordination_v))
g <- ggplot(data=coordination_df, aes(x=coordination_v, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("Strength of aggreement") + theme(axis.text.x = element_blank())  + ggtitle("National results"), vp = viewport(x=0, y=0, width = 0.6, height = 1, just = c('left', 'bottom')))

coordination_v_utility <- df_utility$Coordination
coordination_v_utility  <- factor(coordination_v_utility, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
coordination_df_utility <- as.data.frame(table(coordination_v_utility))
g <- ggplot(data=coordination_df_utility, aes(x=coordination_v_utility, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("")  + theme(axis.text.x = element_blank(), axis.text.y = element_blank()) + ggtitle(paste(utility_name, "results")), vp = viewport(x=0.6, y=0, width = 0.4, height = 1, just = c('left', 'bottom')))
```
Fig. X. Frequency of responses by water utility managers to the question: "YYY," given a rating scale between 'strongly disagree' to 'strongly agree.'

### The effectiveness of adaptation is appropriately monitored and reviewed at my water utility/organisation

```{r fig.width=7, fig.height=2, echo=FALSE} 
grid.newpage()
monitored_v <- df$Monitored
monitored_v  <- factor(monitored_v, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
monitored_df <- as.data.frame(table(monitored_v))
g <- ggplot(data=monitored_df, aes(x=monitored_v, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("Strength of aggreement") + theme(axis.text.x = element_blank())  + ggtitle("National results"), vp = viewport(x=0, y=0, width = 0.6, height = 1, just = c('left', 'bottom')))

monitored_v_utility <- df_utility$Monitored
monitored_v_utility  <- factor(monitored_v_utility, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
monitored_df_utility <- as.data.frame(table(monitored_v_utility))
g <- ggplot(data=monitored_df_utility, aes(x=monitored_v_utility, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("") + theme(axis.text.x = element_blank(), axis.text.y = element_blank()) + ggtitle(paste(utility_name, "results")), vp = viewport(x=0.6, y=0, width = 0.4, height = 1, just = c('left', 'bottom')))
```
Fig. X. Frequency of responses by water utility managers to the question: "YYY," given a rating scale between 'strongly disagree' to 'strongly agree.'

### The internal and external communication and engagement pathways at our organisation are effective for climate adaptation work

```{r fig.width=7, fig.height=2, echo=FALSE} 
grid.newpage()
communication_v <- df$Communication
communication_v  <- factor(communication_v, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
communication_df <- as.data.frame(table(communication_v))
g <- ggplot(data=communication_df, aes(x=communication_v, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("Strength of aggreement") + theme(axis.text.x = element_blank())  + ggtitle("National results"), vp = viewport(x=0, y=0, width = 0.6, height = 1, just = c('left', 'bottom')))

communication_v_utility <- df_utility$Communication
communication_v_utility  <- factor(communication_v_utility, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
communication_df_utility <- as.data.frame(table(communication_v_utility))
g <- ggplot(data=communication_df_utility, aes(x=communication_v_utility, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("")  + theme(axis.text.x = element_blank(), axis.text.y = element_blank()) + ggtitle(paste(utility_name, "results")), vp = viewport(x=0.6, y=0, width = 0.4, height = 1, just = c('left', 'bottom')))
```
Fig. X. Frequency of responses by water utility managers to the question: "YYY," given a rating scale between 'strongly disagree' to 'strongly agree.'

### Climate change adaptation responsibilities are included within position descriptions and performance contracts

```{r fig.width=7, fig.height=2, echo=FALSE} 
grid.newpage()
positions_v <- df$PositionDescriptions
positions_v  <- factor(positions_v, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
positions_df <- as.data.frame(table(positions_v))
g <- ggplot(data=positions_df, aes(x=positions_v, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("Strength of aggreement") + theme(axis.text.x = element_blank())  + ggtitle("National results"), vp = viewport(x=0, y=0, width = 0.6, height = 1, just = c('left', 'bottom')))

positions_v_utility <- df_utility$PositionDescriptions
positions_v_utility  <- factor(positions_v_utility, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
positions_df_utility <- as.data.frame(table(positions_v_utility))
g <- ggplot(data=positions_df_utility, aes(x=positions_v_utility, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("")  + theme(axis.text.x = element_blank(), axis.text.y = element_blank()) + ggtitle(paste(utility_name, "results")), vp = viewport(x=0.6, y=0, width = 0.4, height = 1, just = c('left', 'bottom')))
```
Fig. X. Frequency of responses by water utility managers to the question: "YYY," given a rating scale between 'strongly disagree' to 'strongly agree.'

### Techniques used for climate adaptation strategy

```{r fig.width=7, fig.height=2, echo=FALSE} 
grid.newpage()
library(reshape2)
techniques_df <- data.frame(Balanced_scorecard = as.vector(table(df$BalancedScoresheet, exclude=NULL)), PEST = as.vector(table(df$PEST, exclude=NULL)), SWOT = as.vector(table(df$SWOT, exclude=NULL)), Gap_analysis = as.vector(table(df$GapAnalysis, exclude=NULL)))
techniques_df <- data.frame(techniques_df, use=c('used', 'unused'))
techniques_df_long <- melt(techniques_df) 
colnames(techniques_df_long)<-c('use', 'technique', 'count')

techniques_long_df <- stack(techniques_df_long, select=)
g <- ggplot(data=techniques_df_long, aes(x=technique, y=count, fill=use))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Balanced_scorecard" = "Balanced scorecard", "PEST" = "PEST analysis",
                              "SWOT" = "SWOT analysis", "Gap_analysis" = "Gap analysis")) + ylab("Proportional use") + xlab("Technique") + theme(axis.text.x = element_blank(), legend.position="none")  + ggtitle("National results"), vp = viewport(x=0, y=0, width = 0.5, height = 1, just = c('left', 'bottom')))

techniques_df_utility <- data.frame(Balanced_scorecard = as.vector(table(df_utility$BalancedScoresheet, exclude=NULL)), PEST = as.vector(table(df_utility$PEST, exclude=NULL)), SWOT = as.vector(table(df_utility$SWOT, exclude=NULL)), Gap_analysis = as.vector(table(df_utility$GapAnalysis, exclude=NULL)))
techniques_df_utility <- data.frame(techniques_df_utility, use=c('used', 'unused'))
techniques_df_utility_long <- melt(techniques_df_utility) 
colnames(techniques_df_utility_long)<-c('use', 'technique', 'count')

techniques_long_df_utility <- stack(techniques_df_utility_long, select=)
g <- ggplot(data=techniques_df_utility_long, aes(x=technique, y=count, fill=use))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Balanced_scorecard" = "Balanced scorecard", "PEST" = "PEST analysis",
                              "SWOT" = "SWOT analysis", "Gap_analysis" = "Gap analysis")) + ylab("Proportional use") + xlab("")  + theme(axis.text.x = element_blank(), axis.text.y = element_blank()) + ggtitle(paste(utility_name, "results")), vp = viewport(x=0.5, y=0, width = 0.5, height = 1, just = c('left', 'bottom')))
```
Fig. X. Frequency of responses by water utility managers to the question: "YYY," given a rating scale between 'strongly disagree' to 'strongly agree.'

#### Other strategies used

```{r echo=FALSE, results='hide'}
responses_used <- df_other_utilities$OtherStrategiesUsed
responses_should <- df_other_utilities$OtherStrategiesUsed
responses_used_utility <- df_utility$OtherStrategiesUsed
responses_should_utility <- df_utility$OtherStrategiesUsed
```

Responses from `r utility_name` were:

```{r echo=FALSE, results='asis'}
cat(paste('-',responses_used_utility[!is.na(responses_used_utility)]), sep="\n")
```

While other responses from other Australian utilities, included:

```{r echo=FALSE, results='asis'}
cat(paste('-',responses_used[!is.na(responses_used)]), sep="\n")
```

#### Other strategies that should be used

Responses from `r utility_name` were:

```{r echo=FALSE, results='asis'}
cat(paste('-',responses_should_utility[!is.na(responses_should_utility)]), sep="\n")
```

While other responses from other Australian utilities, included:

```{r echo=FALSE, results='asis'}
cat(paste('-',responses_should[!is.na(responses_should)]), sep="\n")
```

### The use of scenario planning for developing corporate strategy related to climate change adaptation
```{r fig.width=7, fig.height=2, echo=FALSE} 
grid.newpage()
scenarios_v <- df$ScenarioPlanning
scenarios_v  <- factor(scenarios_v, levels=c('No', 'Yes'), ordered = TRUE)
scenarios_df <- as.data.frame(table(scenarios_v))
scenarios_df <- data.frame(scenarios_df, dummy = c(1, 1))
g <- ggplot(data=scenarios_df, aes(x=dummy, y=Freq, fill=scenarios_v))
# http://www.sthda.com/english/wiki/ggplot2-pie-chart-quick-start-guide-r-software-and-data-visualization 
print(g + geom_col() +  coord_flip() + ylab("Proprtional use") + theme(axis.text.x = element_blank(), axis.ticks = element_blank(), axis.text.y = element_blank(), legend.position="none") + xlab("") + coord_polar("y", start=0) + ggtitle("National results"), vp = viewport(x=0, y=0, width = 0.5, height = 1, just = c('left', 'bottom')))
      
scenarios_v_utility <- df_utility$ScenarioPlanning
scenarios_v_utility  <- factor(scenarios_v_utility, levels=c('No', 'Yes'), ordered = TRUE)
scenarios_df_utility <- as.data.frame(table(scenarios_v_utility))
scenarios_df_utility <- data.frame(scenarios_df_utility, dummy = c(1, 1))
g <- ggplot(data=scenarios_df_utility, aes(x=dummy, y=Freq, fill=scenarios_v_utility))
# http://www.sthda.com/english/wiki/ggplot2-pie-chart-quick-start-guide-r-software-and-data-visualization 
print(g + geom_col() +  coord_flip() + ylab("Proprtional use") + theme(axis.text.x = element_blank(), axis.ticks = element_blank(), axis.text.y = element_blank()) + xlab("") + coord_polar("y", start=0) + ggtitle(paste(utility_name, "results")), vp = viewport(x=0.5, y=0, width = 0.5, height = 1, just = c('left', 'bottom'))) 
```
Fig. X. Frequency of responses by water utility managers to the question: "YYY," given a rating scale between 'strongly disagree' to 'strongly agree.'

### Scenario planning is a core method used to ensure flexible long-term plans to create a resilient utiity that is robust with respect to external forces, including climate change

```{r fig.width=7, fig.height=2, echo=FALSE} 
grid.newpage()
sc_extensiveness_v <- df$ScenarioPlanningExtensiveness
sc_extensiveness_v  <- factor(sc_extensiveness_v, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
sc_extensiveness_df <- as.data.frame(table(sc_extensiveness_v))
g <- ggplot(data=sc_extensiveness_df, aes(x=sc_extensiveness_v, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("Strength of aggreement") + theme(axis.text.x = element_blank())  + ggtitle("National results"), vp = viewport(x=0, y=0, width = 0.6, height = 1, just = c('left', 'bottom')))

sc_extensiveness_v_utility <- df_utility$ScenarioPlanningExtensiveness
sc_extensiveness_v_utility  <- factor(sc_extensiveness_v_utility, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
sc_extensiveness_df_utility <- as.data.frame(table(sc_extensiveness_v_utility))
g <- ggplot(data=sc_extensiveness_df_utility, aes(x=sc_extensiveness_v_utility, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("")  + theme(axis.text.x = element_blank(), axis.text.y = element_blank()) + ggtitle(paste(utility_name, "results")), vp = viewport(x=0.6, y=0, width = 0.4, height = 1, just = c('left', 'bottom')))
```
Fig. X. Frequency of responses by water utility managers to the question: "YYY," given a rating scale between 'strongly disagree' to 'strongly agree.'

### Consistent sets of climate change scenarios are used across entire organisations

```{r fig.width=7, fig.height=2, echo=FALSE} 
grid.newpage()
consistent_sc_v <- df$ConsistentScenarios
consistent_sc_v  <- factor(consistent_sc_v, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
consistent_sc_df <- as.data.frame(table(consistent_sc_v))
g <- ggplot(data=consistent_sc_df, aes(x=consistent_sc_v, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("Strength of aggreement") + theme(axis.text.x = element_blank())  + ggtitle("National results"), vp = viewport(x=0, y=0, width = 0.6, height = 1, just = c('left', 'bottom')))

consistent_sc_v_utility <- df_utility$ConsistentScenarios
consistent_sc_v_utility  <- factor(consistent_sc_v_utility, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
consistent_sc_df_utility <- as.data.frame(table(consistent_sc_v_utility))
g <- ggplot(data=consistent_sc_df_utility, aes(x=consistent_sc_v_utility, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("")  + theme(axis.text.x = element_blank(), axis.text.y = element_blank()) + ggtitle(paste(utility_name, "results")), vp = viewport(x=0.6, y=0, width = 0.4, height = 1, just = c('left', 'bottom'))) 
```
Fig. X. Frequency of responses by water utility managers to the question: "YYY," given a rating scale between 'strongly disagree' to 'strongly agree.'

### Scenarios include regulatory and innovation factors
```{r fig.width=7, fig.height=4, echo=FALSE} 
grid.newpage()
regulatory_sc_v <- df$RegulatoryScenarios
innovation_sc_v <- df$InnovationScenarios
regulatory_sc_v  <- factor(regulatory_sc_v, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
innovation_sc_v  <- factor(innovation_sc_v, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
sc_factors_df <- data.frame(as.data.frame(table(regulatory_sc_v)), as.data.frame(table(innovation_sc_v))[2])
names(sc_factors_df) <- c('response', 'regulatory factors', 'innovation factors')
sc_factors_df <- melt(sc_factors_df)
g <- ggplot(data=sc_factors_df, aes(x=response, y=value))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("Strength of aggreement") + theme(axis.text.x = element_blank())  + ggtitle("National results") + facet_grid(variable ~ .), vp = viewport(x=0, y=0, width = 0.6, height = 1, just = c('left', 'bottom')))

regulatory_sc_v_utility <- df_utility$RegulatoryScenarios
innovation_sc_v_utility <- df_utility$InnovationScenarios
regulatory_sc_v_utility  <- factor(regulatory_sc_v_utility, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
innovation_sc_v_utility  <- factor(innovation_sc_v_utility, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
sc_factors_df_utility <- data.frame(as.data.frame(table(regulatory_sc_v_utility)), as.data.frame(table(innovation_sc_v_utility))[2])
names(sc_factors_df_utility) <- c('response', 'regulatory factors', 'innovation factors')
sc_factors_df_utility <- melt(sc_factors_df_utility)
g <- ggplot(data=sc_factors_df_utility, aes(x=response, y=value))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("")  + theme(axis.text.x = element_blank(), axis.text.y = element_blank()) + ggtitle(paste(utility_name, "results")) + facet_grid(variable ~ .), vp = viewport(x=0.6, y=0, width = 0.4, height = 1, just = c('left', 'bottom')))
```
Fig. X. Frequency of responses by water utility managers to the question: "YYY," given a rating scale between 'strongly disagree' to 'strongly agree.'

### Possible, relevant and challenging scenarios are used to help identify key weaknesses and opportunities

```{r fig.width=7, fig.height=2, echo=FALSE} 
grid.newpage()
possible_challenging_sc_v <- df$PossibleRelevantChallengingScenarios
possible_challenging_sc_v  <- factor(possible_challenging_sc_v, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
possible_challenging_sc_df <- as.data.frame(table(possible_challenging_sc_v))
g <- ggplot(data=possible_challenging_sc_df, aes(x=possible_challenging_sc_v, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("Strength of aggreement") + theme(axis.text.x = element_blank())  + ggtitle("National results"), vp = viewport(x=0, y=0, width = 0.6, height = 1, just = c('left', 'bottom')))

possible_challenging_sc_v_utility <- df_utility$PossibleRelevantChallengingScenarios
possible_challenging_sc_v_utility  <- factor(possible_challenging_sc_v_utility, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
possible_challenging_sc_df_utility <- as.data.frame(table(possible_challenging_sc_v_utility))
g <- ggplot(data=possible_challenging_sc_df_utility, aes(x=possible_challenging_sc_v_utility, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("")  + theme(axis.text.x = element_blank(), axis.text.y = element_blank()) + ggtitle(paste(utility_name, "results")), vp = viewport(x=0.6, y=0, width = 0.4, height = 1, just = c('left', 'bottom')))
```
Fig. X. Frequency of responses by water utility managers to the question: "YYY," given a rating scale between 'strongly disagree' to 'strongly agree.'

### Human influence on the climate system is clear, and recent anthropogenic emissions of greenhouse gases are the highest in history

```{r fig.width=7, fig.height=2, echo=FALSE} 
grid.newpage()
anthropogenic_change_clear_v <- df$OpinionClimaticHumanInfluence
anthropogenic_change_clear_v  <- factor(anthropogenic_change_clear_v, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
anthropogenic_change_clear_df <- as.data.frame(table(anthropogenic_change_clear_v))
g <- ggplot(data=anthropogenic_change_clear_df, aes(x=anthropogenic_change_clear_v, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("Strength of aggreement") + theme(axis.text.x = element_blank())  + ggtitle("National results"), vp = viewport(x=0, y=0, width = 0.6, height = 1, just = c('left', 'bottom')))

anthropogenic_change_clear_v_utility <- df_utility$OpinionClimaticHumanInfluence
anthropogenic_change_clear_v_utility  <- factor(anthropogenic_change_clear_v_utility, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
anthropogenic_change_clear_df_utility <- as.data.frame(table(anthropogenic_change_clear_v_utility))
g <- ggplot(data=anthropogenic_change_clear_df_utility, aes(x=anthropogenic_change_clear_v_utility, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("")  + theme(axis.text.x = element_blank(), axis.text.y = element_blank()) + ggtitle(paste(utility_name, "results")), vp = viewport(x=0.6, y=0, width = 0.4, height = 1, just = c('left', 'bottom')))
```
Fig. X. Frequency of responses by water utility managers to the question: "YYY," given a rating scale between 'strongly disagree' to 'strongly agree.'

### Recent climate changes have already had widespread impacts on human and natural systems

```{r fig.width=7, fig.height=2, echo=FALSE} 
grid.newpage()
anthropogenic_change_already_v <- df$OpinionPresentHumanImpact
anthropogenic_change_already_v  <- factor(anthropogenic_change_already_v, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
anthropogenic_change_already_df <- as.data.frame(table(anthropogenic_change_already_v))
g <- ggplot(data=anthropogenic_change_already_df, aes(x=anthropogenic_change_already_v, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("Strength of aggreement") + theme(axis.text.x = element_blank())  + ggtitle("National results"), vp = viewport(x=0, y=0, width = 0.6, height = 1, just = c('left', 'bottom')))

anthropogenic_change_already_v_utility <- df_utility$OpinionPresentHumanImpact
anthropogenic_change_already_v_utility  <- factor(anthropogenic_change_already_v_utility, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
anthropogenic_change_already_df_utility <- as.data.frame(table(anthropogenic_change_already_v_utility))
g <- ggplot(data=anthropogenic_change_already_df_utility, aes(x=anthropogenic_change_already_v_utility, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("")  + theme(axis.text.x = element_blank(), axis.text.y = element_blank()) + ggtitle(paste(utility_name, "results")), vp = viewport(x=0.6, y=0, width = 0.4, height = 1, just = c('left', 'bottom')))
```
Fig. X. Frequency of responses by water utility managers to the question: "YYY," given a rating scale between 'strongly disagree' to 'strongly agree.'

### The science is clear that continued emission of greenhouse gases will cause further warming and long-lasting changes in all components of the climate system, increasing the likelihood of severe, pervasive and irreversible impacts for people and ecosystems

```{r fig.width=7, fig.height=2, echo=FALSE} 
grid.newpage()
science_clear_v <- df$OpinionScienceClear
science_clear_v  <- factor(science_clear_v, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
science_clear_df <- as.data.frame(table(science_clear_v))
g <- ggplot(data=science_clear_df, aes(x=science_clear_v, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("Strength of aggreement") + theme(axis.text.x = element_blank())  + ggtitle("National results"), vp = viewport(x=0, y=0, width = 0.6, height = 1, just = c('left', 'bottom')))

science_clear_v_utility <- df_utility$OpinionScienceClear
science_clear_v_utility  <- factor(science_clear_v_utility, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
science_clear_df_utility <- as.data.frame(table(science_clear_v_utility))
g <- ggplot(data=science_clear_df_utility, aes(x=science_clear_v_utility, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("")  + theme(axis.text.x = element_blank(), axis.text.y = element_blank()) + ggtitle(paste(utility_name, "results")), vp = viewport(x=0.6, y=0, width = 0.4, height = 1, just = c('left', 'bottom'))) 
```
Fig. X. Frequency of responses by water utility managers to the question: "YYY," given a rating scale between 'strongly disagree' to 'strongly agree.'

### Climate change significantly threatens??the reliability of water supply to consumers

```{r fig.width=7, fig.height=2, echo=FALSE} 
grid.newpage()
reliability_v <- df$OpinionClimateThreatensReliability
reliability_v  <- factor(reliability_v, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
reliability_df <- as.data.frame(table(reliability_v))
g <- ggplot(data=reliability_df, aes(x=reliability_v, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("Strength of aggreement") + theme(axis.text.x = element_blank())  + ggtitle("National results"), vp = viewport(x=0, y=0, width = 0.6, height = 1, just = c('left', 'bottom')))

reliability_v_utility <- df_utility$OpinionClimateThreatensReliability
reliability_v_utility  <- factor(reliability_v_utility, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
reliability_df_utility <- as.data.frame(table(reliability_v_utility))
g <- ggplot(data=reliability_df_utility, aes(x=reliability_v_utility, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("Strength of aggreement")  + theme(axis.text.x = element_blank(), axis.text.y = element_blank()) + ggtitle(paste(utility_name, "results")), vp = viewport(x=0.6, y=0, width = 0.4, height = 1, just = c('left', 'bottom')))
```
Fig. X. Frequency of responses by water utility managers to the question: "YYY," given a rating scale between 'strongly disagree' to 'strongly agree.'

### Climate change significantly threatens??the affordability of water supply to consumers

```{r fig.width=7, fig.height=2, echo=FALSE} 
grid.newpage()
affordability_v <- df$OpinionClimateThreatensReliability
affordability_v  <- factor(affordability_v, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
affordability_df <- as.data.frame(table(affordability_v))
g <- ggplot(data=affordability_df, aes(x=affordability_v, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("Strength of aggreement") + theme(axis.text.x = element_blank())  + ggtitle("National results"), vp = viewport(x=0, y=0, width = 0.6, height = 1, just = c('left', 'bottom')))

affordability_v_utility <- df_utility$OpinionClimateThreatensReliability
affordability_v_utility  <- factor(affordability_v_utility, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
affordability_df_utility <- as.data.frame(table(affordability_v_utility))
g <- ggplot(data=affordability_df_utility, aes(x=affordability_v_utility, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("Strength of aggreement")  + theme(axis.text.x = element_blank(), axis.text.y = element_blank()) + ggtitle(paste(utility_name, "results")), vp = viewport(x=0.6, y=0, width = 0.4, height = 1, just = c('left', 'bottom')))
```
Fig. X. Frequency of responses by water utility managers to the question: "YYY," given a rating scale between 'strongly disagree' to 'strongly agree.'

### Climate change will significantly affect the consumptive behaviours or water??demands of consumers

```{r fig.width=7, fig.height=2, echo=FALSE} 
grid.newpage()
behaviour_v <- df$OpinionClimateAffectrsBehaviour
behaviour_v  <- factor(behaviour_v, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
behaviour_df <- as.data.frame(table(behaviour_v))
g <- ggplot(data=behaviour_df, aes(x=behaviour_v, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("Strength of aggreement") + theme(axis.text.x = element_blank())  + ggtitle("National results"), vp = viewport(x=0, y=0, width = 0.6, height = 1, just = c('left', 'bottom')))

behaviour_v_utility <- df_utility$OpinionClimateAffectrsBehaviour
behaviour_v_utility  <- factor(behaviour_v_utility, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
behaviour_df_utility <- as.data.frame(table(behaviour_v_utility))
g <- ggplot(data=behaviour_df_utility, aes(x=behaviour_v_utility, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("") + theme(axis.text.x = element_blank(), axis.text.y = element_blank()) + ggtitle(paste(utility_name, "results")), vp = viewport(x=0.6, y=0, width = 0.4, height = 1, just = c('left', 'bottom')))
```
Fig. X. Frequency of responses by water utility managers to the question: "YYY," given a rating scale between 'strongly disagree' to 'strongly agree.'

### Climate change will significantly affect??business processes

```{r fig.width=7, fig.height=2, echo=FALSE} 
grid.newpage()
business_v <- df$OpinionClimateAffectrsBusiness
business_v  <- factor(business_v, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
business_df <- as.data.frame(table(business_v))
g <- ggplot(data=business_df, aes(x=business_v, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("Strength of aggreement") + theme(axis.text.x = element_blank())  + ggtitle("National results"), vp = viewport(x=0, y=0, width = 0.6, height = 1, just = c('left', 'bottom')))

business_v_utility <- df_utility$OpinionClimateAffectrsBusiness
business_v_utility  <- factor(business_v_utility, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
business_df_utility <- as.data.frame(table(business_v_utility))
g <- ggplot(data=business_df_utility, aes(x=business_v_utility, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("")  + theme(axis.text.x = element_blank(), axis.text.y = element_blank()) + ggtitle(paste(utility_name, "results")), vp = viewport(x=0.6, y=0, width = 0.4, height = 1, just = c('left', 'bottom')))
```
Fig. X. Frequency of responses by water utility managers to the question: "YYY," given a rating scale between 'strongly disagree' to 'strongly agree.'

### For the water sector, enhancing capacity to mitigate and adapt are part of the essential foundation for managing climate change risks

```{r fig.width=7, fig.height=2, echo=FALSE} 
grid.newpage()
foundation_v <- df$OpinionAdaptationFoundation
foundation_v  <- factor(foundation_v, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
foundation_df <- as.data.frame(table(foundation_v))
g <- ggplot(data=foundation_df, aes(x=foundation_v, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("Strength of aggreement") + theme(axis.text.x = element_blank())  + ggtitle("National results"), vp = viewport(x=0, y=0, width = 0.6, height = 1, just = c('left', 'bottom')))

foundation_v_utility <- df_utility$OpinionAdaptationFoundation
foundation_v_utility  <- factor(foundation_v_utility, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
foundation_df_utility <- as.data.frame(table(foundation_v_utility))
g <- ggplot(data=foundation_df_utility, aes(x=foundation_v_utility, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("")  + theme(axis.text.x = element_blank(), axis.text.y = element_blank()) + ggtitle(paste(utility_name, "results")), vp = viewport(x=0.6, y=0, width = 0.4, height = 1, just = c('left', 'bottom')))
```
Fig. X. Frequency of responses by water utility managers to the question: "YYY," given a rating scale between 'strongly disagree' to 'strongly agree.'

### Many adaptation and mitigation options can help address climate change, but no single option is sufficient in??itself.

```{r fig.width=7, fig.height=2, echo=FALSE} 
grid.newpage()
portfolios_v <- df$OpinionCombinatorialApproach
portfolios_v  <- factor(portfolios_v, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
portfolios_df <- as.data.frame(table(portfolios_v))
g <- ggplot(data=portfolios_df, aes(x=portfolios_v, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("Strength of aggreement") + theme(axis.text.x = element_blank())  + ggtitle("National results"), vp = viewport(x=0, y=0, width = 0.6, height = 1, just = c('left', 'bottom')))

portfolios_v_utility <- df_utility$OpinionCombinatorialApproach
portfolios_v_utility  <- factor(portfolios_v_utility, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
portfolios_df_utility <- as.data.frame(table(portfolios_v_utility))
g <- ggplot(data=portfolios_df_utility, aes(x=portfolios_v_utility, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("") + theme(axis.text.x = element_blank(), axis.text.y = element_blank()) + ggtitle(paste(utility_name, "results")), vp = viewport(x=0.6, y=0, width = 0.4, height = 1, just = c('left', 'bottom')))
```
Fig. X. Frequency of responses by water utility managers to the question: "YYY," given a rating scale between 'strongly disagree' to 'strongly agree.'

#### Comments
```{r echo=FALSE, results='hide'}
responses <- df_other_utilities$CommentsCombinatorialApproach
responses <- responses[!is.na(responses)]
responses_utility <- df_utility$CommentsCombinatorialApproach
responses_utility <- responses_utility[!is.na(responses_utility)]
```

Comments from `r utility_name` were:
```{r echo=FALSE, results='asis'}
cat(paste('-',responses_utility), sep="\n")
```

While other comments from other Australian utilities, included:
```{r echo=FALSE, results='asis'}
cat(paste('-',responses), sep="\n")
```


### Effective implementation depends on policies and cooperation at all scales and can be enhanced through integrated responses that link mitigation and adaptation with other water utility objectives/KPIs.

```{r fig.width=7, fig.height=2, echo=FALSE} 
grid.newpage()
holistic_integrated_v <- df$OpinionIntegrationCooperation
holistic_integrated_v  <- factor(holistic_integrated_v, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
holistic_integrated_df <- as.data.frame(table(holistic_integrated_v))
g <- ggplot(data=holistic_integrated_df, aes(x=holistic_integrated_v, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("Strength of aggreement") + theme(axis.text.x = element_blank())  + ggtitle("National results"), vp = viewport(x=0, y=0, width = 0.6, height = 1, just = c('left', 'bottom')))

holistic_integrated_v_utility <- df_utility$OpinionIntegrationCooperation
holistic_integrated_v_utility <- factor(holistic_integrated_v_utility, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
holistic_integrated_df_utility <- as.data.frame(table(holistic_integrated_v_utility))
g <- ggplot(data=holistic_integrated_df_utility, aes(x=holistic_integrated_v_utility, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("")  + theme(axis.text.x = element_blank(), axis.text.y = element_blank()) + ggtitle(paste(utility_name, "results")), vp = viewport(x=0.6, y=0, width = 0.4, height = 1, just = c('left', 'bottom')))
```
Fig. X. Frequency of responses by water utility managers to the question: "YYY," given a rating scale between 'strongly disagree' to 'strongly agree.'

### Adaptive water management techniques, including scenario planning, learning-based approaches and flexible and low-regret solutions, are helpful in adapting to uncertain hydrological changes

```{r fig.width=7, fig.height=2, echo=FALSE} 
grid.newpage()
helpful_approaches_v <- df$OpinionHelpfulApproaches
helpful_approaches_v  <- factor(helpful_approaches_v, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
helpful_approaches_df <- as.data.frame(table(helpful_approaches_v))
g <- ggplot(data=helpful_approaches_df, aes(x=helpful_approaches_v, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("Strength of aggreement") + theme(axis.text.x = element_blank())  + ggtitle("National results"), vp = viewport(x=0, y=0, width = 0.6, height = 1, just = c('left', 'bottom')))

helpful_approaches_v_utility <- df_utility$OpinionHelpfulApproaches
helpful_approaches_v_utility  <- factor(helpful_approaches_v_utility, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
helpful_approaches_df_utility <- as.data.frame(table(helpful_approaches_v_utility))
g <- ggplot(data=helpful_approaches_df_utility, aes(x=helpful_approaches_v_utility, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("") + theme(axis.text.x = element_blank(), axis.text.y = element_blank()) + ggtitle(paste(utility_name, "results")), vp = viewport(x=0.6, y=0, width = 0.4, height = 1, just = c('left', 'bottom')))
```
Fig. X. Frequency of responses by water utility managers to the question: "YYY," given a rating scale between 'strongly disagree' to 'strongly agree.'

### The integration of adaptation into planning and decision-making plays a key role in translating climate change adaptation planning to implementation

```{r fig.width=7, fig.height=2, echo=FALSE}
grid.newpage()
integration_v <- df$OpinionAdaptationInPlanningDecisionMaking
integration_v  <- factor(integration_v, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
integration__df <- as.data.frame(table(integration_v))
g <- ggplot(data=integration__df, aes(x=integration_v, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("Strength of aggreement") + theme(axis.text.x = element_blank())  + ggtitle("National results"), vp = viewport(x=0, y=0, width = 0.6, height = 1, just = c('left', 'bottom')))

integration_v_utility <- df_utility$OpinionAdaptationInPlanningDecisionMaking
integration_v_utility  <- factor(integration_v_utility, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
integration__df_utility <- as.data.frame(table(integration_v_utility))
g <- ggplot(data=integration__df_utility, aes(x=integration_v_utility, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("")  + theme(axis.text.x = element_blank(), axis.text.y = element_blank()) + ggtitle(paste(utility_name, "results")), vp = viewport(x=0.6, y=0, width = 0.4, height = 1, just = c('left', 'bottom')))
```
Fig. X. Frequency of responses by water utility managers to the question: "YYY," given a rating scale between 'strongly disagree' to 'strongly agree.'

### Investment in research and diffusion/transfer of research in support of climate change adaptation is important

```{r fig.width=7, fig.height=2, echo=FALSE} 
grid.newpage()
research_v <- df$OpinionResearchImportant
research_v  <- factor(research_v, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
research_df <- as.data.frame(table(research_v))
g <- ggplot(data=research_df, aes(x=research_v, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("Strength of aggreement") + theme(axis.text.x = element_blank())  + ggtitle("National results"), vp = viewport(x=0, y=0, width = 0.6, height = 1, just = c('left', 'bottom')))

research_v_utility <- df_utility$OpinionResearchImportant
research_v_utility  <- factor(research_v_utility, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
research_df_utility <- as.data.frame(table(research_v_utility))
g <- ggplot(data=research_df_utility, aes(x=research_v_utility, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("")  + theme(axis.text.x = element_blank(), axis.text.y = element_blank()) + ggtitle(paste(utility_name, "results")), vp = viewport(x=0.6, y=0, width = 0.4, height = 1, just = c('left', 'bottom')))
```
Fig. X. Frequency of responses by water utility managers to the question: "YYY," given a rating scale between 'strongly disagree' to 'strongly agree.'

### Effective climate change adaptation involves an integrated approach with other service providers and managers of public goods (e.g. energy, telecommunications, transport, supply chains)

```{r fig.width=7, fig.height=2, echo=FALSE} 
grid.newpage()
interdependencies_v <- df$OpinionInterdependencies
interdependencies_v  <- factor(interdependencies_v, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
interdependencies_df <- as.data.frame(table(interdependencies_v))
g <- ggplot(data=interdependencies_df, aes(x=interdependencies_v, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("Strength of aggreement") + theme(axis.text.x = element_blank())  + ggtitle("National results"), vp = viewport(x=0, y=0, width = 0.6, height = 1, just = c('left', 'bottom')))

interdependencies_v_utility <- df_utility$OpinionInterdependencies
interdependencies_v_utility  <- factor(interdependencies_v_utility, levels=c('Not sure', 'Strongly disagree', '2', 'Neither agree nor disagree', '4', 'Strongly agree'), ordered = TRUE)
interdependencies_df_utility <- as.data.frame(table(interdependencies_v_utility))
g <- ggplot(data=interdependencies_df_utility, aes(x=interdependencies_v_utility, y=Freq))
print(g + geom_col() +  coord_flip() + scale_x_discrete(labels=c("Not sure" = "Not sure", "Strongly disagree" = "Strongly disagree",
                              "2" = "", "Neither agree nor disagree" = "Neither agree nor disagree", "4" = "", "Strongly agree" =  "Strongly agree")) + ylab("Frequency of response") + xlab("")  + theme(axis.text.x = element_blank(), axis.text.y = element_blank()) + ggtitle(paste(utility_name, "results")), vp = viewport(x=0.6, y=0, width = 0.4, height = 1, just = c('left', 'bottom')))
```
Fig. X. Frequency of responses by water utility managers to the question: "YYY," given a rating scale between 'strongly disagree' to 'strongly agree.'

### Core areas surveyed

```{r fig.width=7, fig.height=2, echo=FALSE} 
grid.newpage()
core_areas_v <- df$CoreArea
core_areas_v  <- factor(core_areas_v, levels=c('Source warers and supply', 'Built assets', 'Environmental management', 'People and Workplaces', 'Customer and product delivery', 'Corporate governance'), ordered = FALSE)
core_areas_df <- as.data.frame(table(core_areas_v))
g <- ggplot(data=core_areas_df, aes(x=core_areas_v, y=Freq))
print(g + geom_col() +  coord_flip() + ylab("Frequency of response") + xlab("Core areas") + theme(axis.text.x = element_blank())  + ggtitle("National results"), vp = viewport(x=0, y=0, width = 0.6, height = 1, just = c('left', 'bottom')))

core_areas_v_utility <- df_utility$CoreArea
core_areas_v_utility  <- factor(core_areas_v_utility, levels=c('Source warers and supply', 'Built assets', 'Environmental management', 'People and Workplaces', 'Customer and product delivery', 'Corporate governance'), ordered = FALSE)
core_areas_df_utility <- as.data.frame(table(core_areas_v_utility))
g <- ggplot(data=core_areas_df_utility, aes(x=core_areas_v_utility, y=Freq))
print(g + geom_col() +  coord_flip() + ylab("Frequency of response") + xlab("") + theme(axis.text.x = element_blank(), axis.text.y = element_blank()) + ggtitle(paste(utility_name, "results")), vp = viewport(x=0.6, y=0, width = 0.4, height = 1, just = c('left', 'bottom')))
```
Fig. X. Frequency of responses by water utility managers to the question: "YYY," given a rating scale between 'strongly disagree' to 'strongly agree.'

